---
output:
  pdf_document:
    highlight: tango
  html_document: default
---

 Exponential Distribution Simulation and Demonstrate Central Limit Theorem
---
 Author: Ziwen Yu  
 Date: `r Sys.Date()`  
--------------------------------------------------------------------------------------------

This report, generated in **knitr**, contains a simulation on exponential distribution and a demonstratation of the "Central Limit Theorem". The analysis is  performed in **R** and uses **ggplot2** for plotting.
The rate of exponential distribution which the random numbers are generated from is **$\lambda=0.2$**.  


**Step 1**:  

```{r}
set.seed(10)      # set seed for generating random numbers
rate=0.2          # Rate of exponential distribution
n=40              # Number of samples for each sample set
N_rep=1000        # Replication times

# Generate 40 exponentially distributed random numbers for a sample set
# Calculate the mean of each sample set
# Replicate this process 1000 times
Dt=data.frame(Means=replicate(N_rep,mean(rexp(n,rate))))

summary(Dt$Means)
```

**Step 2**:  

It could be seen from the following code that the sample mean (5.04506) is very close to the theoretical mean of the population ($\lambda^{-1}$= `r 1/rate`). 
```{r}

1/rate            # Theoretical mean
mean(Dt$Means)    # Sample mean

```
The sample variance (0.6372544) is also very close to the theoretical variance which is $\sigma^2 / n = 1/(\lambda^2 n)$ =`r 1/(0.04 * 40)`.
```{r}

1/rate^2/n        # Theoretical variance)
sd(Dt$Means)^2  # Sample variance
```

The comparison between sample distribution and theoretical distribution is provided below. 
```{r,warning=FALSE,message=FALSE,fig.height=4,fig.width=6}
# Load plotting package and defining plot theme
library(ggplot2)
Plot_theme = theme_bw()+
      theme(plot.title = element_text(size = 12)
            ,axis.title.x = element_text(size=10)
            ,axis.title.y = element_text(size=10)
            ,axis.text.x = element_text(size=8)
            ,axis.text.y = element_text(size=8)            
      )

x=seq(from=3,to=7,by=0.1)
DT1=data.frame(x=x,y=dnorm(x,mean=1/rate,sd=1/rate/sqrt(n)))

# Plot the distribution of sample set means
ggplot(data=Dt,aes(Means))+
      geom_histogram(aes(y=..density..),fill='grey',binwidth=0.1)+
      stat_density(col='black',fill=NA,size=1)+
      geom_line(data=DT1,aes(x,y),col='red',position = "stack",size=1,linetype="dashed")+
      geom_vline(xintercept=1/rate,col='red',size=1,linetype="dashed")+
      geom_vline(xintercept=mean(Dt$Means),col='black',size=1)+
      Plot_theme+
      labs(title="Sample Distribution vs. Theoretical Distribution")
```

**Step 3**:  

A Quantile-Quantile (QQ) plot is employed to further test its normality. As the following figure shows, the distribution of sample set means overlaps with the normality line around 5 (its mean). The tails do not match normality very well due to the small size of each sample set. 

```{r,fig.height=4,fig.width=6}
 
# Code copied from qqline() in R to plot the normality line in QQ plot using ggplot2
gg_qqline<-function(y,probs = c(0.25,0.75)) {
      distribution=qnorm
      stopifnot(length(probs) == 2, is.function(distribution))
      y <- quantile(y, probs, names = FALSE, type = 7, na.rm = TRUE)
      x <- distribution(probs)
      slope <- diff(y)/diff(x)
      int <- y[1L] - slope * x[1L]
      geom_abline(slope=slope,intercept =int, linetype="dashed",col='red',size=1)
}

# QQ plot using ggplot2
ggplot(data=Dt,aes(sample=Means))+
      stat_qq()+
      gg_qqline(Dt$Means)+
      Plot_theme+
      labs(title='Normal Probability Plot'
           ,x='Theoretical Normal Distribution'
           ,y='Sample Set Mean Distribution')
```

The final analysis is to evaluate the coverage of the confidence interval for
$1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$
```{r}
ll <- Dt$Means - qnorm(0.975) * sqrt(1/rate^2/n)
ul <- Dt$Means + qnorm(0.975) * sqrt(1/rate^2/n)
CI=mean(ll < 1/rate & ul > 1/rate)
```
The confidence interval for the coverage of theoretical mean is **`r CI*100`**%.

