---
title: "Data Scientist Capstone Project Milestone Report"
author: "Ziwen Yu"
output: html_document
---

This is the milstone report of the data scientist capstone project on "Yelp Dataset Challenge". In this report, the data will be loaded and explored to get a brief understading of the information. The goal of this analysis is to come out a question that is of interest for some audience and answerable from the data. 

## Loading the data

```{r Load Data,eval=FALSE}

download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/yelp_dataset_challenge_academic_dataset.zip",
              'yelp_dataset_challenge_academic_dataset.zip')
unzip('yelp_dataset_challenge_academic_dataset.zip',
      junkpaths=F)

library(rjson)
library(dplyr)
library(knitr)

# Function to read JSON file
ReadJSON_File = function (fileNM){
      fileNM %>%
            readLines %>%
            lapply(.,fromJSON)
}

business    <- ReadJSON_File('yelp_academic_dataset_business.json')
review      <- ReadJSON_File('yelp_academic_dataset_review.json')
user        <- ReadJSON_File('yelp_academic_dataset_user.json')
checkin     <- ReadJSON_File('yelp_academic_dataset_checkin.json')
tip         <- ReadJSON_File('yelp_academic_dataset_tip.json')
```

```{r Read data from saved files,echo=FALSE,message=FALSE}
library(dplyr)
library(knitr)

setwd('C:\\Users\\Ziwen.Yu\\Documents\\Class\\Capstone\\Data\\Oct 2015\\')

path='Saved Data\\'
business    =readRDS(paste0(path,"business.rds"))
review      =readRDS(paste0(path,"review.rds"))
user        =readRDS(paste0(path,"user.rds"))
checkin     =readRDS(paste0(path,"checkin.rds"))
tip         =readRDS(paste0(path,"tip.rds"))
Health      =readRDS(paste0(path,"Health.rds"))
Sub_cate    =readRDS(paste0(path,"Sub_cate.rds"))
Doc_reviews =readRDS(paste0(path,"Doc_reviews.RDS"))

Obs=c('business',
      'review',
      'user',
      'checkin',
      'tip',
      'Health')

kable(

      cbind(
      sapply(Obs,function(x) format(object.size(get(x)),units='MB')),
      sapply(Obs,function(x) length(get(x)))
      ) %>%
            data.frame() %>%
            rename(Size=X1,Length=X2)
      )

```

## Explore all the business

The question of interest to me is how to detect the business category by analyzing the review content.   __First__ thing to check here is the main business categories in the business data. 

```{r Category Histogram,message=FALSE,error=FALSE}

library(ggplot2)
Plot_theme = theme_bw()+
      theme(plot.title = element_text(size = 12)
            ,axis.title.x = element_text(size=10)
            ,axis.title.y = element_text(size=10)
            ,axis.text.x = element_text(size=10,angle = 45, hjust = 1)
            ,axis.text.y = element_text(size=10)            
      )

category=sapply(business,function(x) x$categories)
category %>%
      unlist() %>%
      data.frame(Categories=.) %>%
      group_by(Categories) %>%
      tally() %>%
      rename(Freq=n) %>%
      top_n(.,20,Freq) %>%
      ggplot(.,aes(reorder(Categories, -Freq), Freq))+
      geom_bar(stat="identity")+
      xlab('')+
      ggtitle('Top 20 Business Categories')+
      Plot_theme

```

__Secondly__, I would focus on the "Health & Medical" category business which is related to my work. The subcategories in "Health & Medical" business are further explored.
```{r SubCategory in Health & medical}
findbusiness_by_Category = function (Category){
      x=sapply(business,function(x) Category %in% x$categories)
      business[x]
}

Category_of_interest='Health & Medical'
Health=findbusiness_by_Category(Category_of_interest)

category=sapply(Health,function(x) x$categories)
category %>%
      unlist() %>%
      data.frame(Categories=.) %>%
      filter(Categories!=Category_of_interest) %>%      
      group_by(Categories) %>%
      tally() %>%
      rename(Freq=n) %>%
      top_n(.,40,Freq) %>%
      ggplot(.,aes(reorder(Categories, -Freq), Freq))+
      geom_bar(stat="identity")+
      xlab('')+
      ggtitle('Top 20 Sub-Categories in Health & Medical Business')+
      Plot_theme

```

## Target Problem 

We can see that "Doctors" is the biggest sub-category in the "Health & Medical" business. Many of the top 20 categories are actually more granular under "Doctors" category, such as "Optometrists", "Family Practice" and "Pediatricians", etc. However, there are still `r sum(sapply(Health,function(x) ('Health & Medical' %in% x$categories) & ('Doctors' %in% x$categories) & length(x$categories)==2))` businesses tagged as "Doctors" with no specialty information.

This could be a problem for those who are looking for a doctors in a specialized field. The target problem of this project is __to build a model that could automatically categorize the specialty of a "Doctors" business based on the existing review information using text mining__. 

## Further Exploratory Analysis

1. Make sure all "Doctors" businesses are under "Health & Medical" category
```{r Check Doctors under Health & Medical}
Category_of_interest='Doctors'
Doctor=findbusiness_by_Category(Category_of_interest)

Doctor %>%
      sapply(.,function(x) x$categories) %>%
      unlist() %>%
      data.frame(Categories=.) %>%
      filter (Categories %in% c('Doctors','Health & Medical')) %>%
      group_by(Categories) %>%
      tally() %>%
      rename(Freq=n) %>%
      kable
```


2. Check the frequency of all sub-categories under "Doctors"
```{r Histogram of sub-categories under Doctors}

Doctor %>%
      sapply(.,function(x) x$categories) %>%
      unlist() %>%
      data.frame(Categories=.) %>%
      filter (Categories !='Doctors' & Categories !='Health & Medical') %>%
      group_by(Categories) %>%
      tally() %>%
      rename(Freq=n) %>%
      top_n(20,Freq) %>%
      ggplot(.,aes(reorder(Categories, -Freq), Freq))+
      geom_bar(stat="identity")+
      xlab('')+
      ggtitle('Top 20 Sub-Categories under Doctors Category')+
      Plot_theme
      
```

3. Check the frequency of the reviews for each sub-categories under "Doctors"
```{r Get review under doctors, eval=FALSE}
## Get all sub_catetories under doctors
Doctor %>%
      sapply(.,function(x) x$categories) %>%
      unlist() %>%
      data.frame(Categories=.) %>%
      filter (Categories !='Doctors' & Categories !='Health & Medical') %>%
      distinct(Categories) ->Sub_cate

# Function to extract business_id under Doctors by a sub-category
business_id_by_Doc_SubCate = function(Category){
      x=sapply(business,function(x) (Category %in% x$categories) &
                     ('Doctors' %in% x$categories) &
                     ('Health & Medical' %in% x$categories))
      SubCate_business=business[x]
      sapply(SubCate_business,function(x) x$business_id)
}

#Get all business IDs for all sub-catetories
Sub_Cate_bus_id=lapply(Sub_cate$Categories,
                       function(x) business_id_by_Doc_SubCate(x))

# Function to get reviews by business ID
findreview_by_business_id = function (ID){
      x=sapply(review,function(x) x$business_id %in% ID)
      review[x]
}

# Get reviews for all business IDs for each sub-category
Doc_reviews=sapply(Sub_Cate_bus_id,function(x) findreview_by_business_id(x))
```

```{r Plot histogram of review frequency}
Sub_cate$ReviewNum=sapply(Doc_reviews,length)

Sub_cate %>%
      top_n(20,ReviewNum) %>%
      ggplot(.,aes(reorder(Categories, -ReviewNum),ReviewNum))+
      geom_bar(stat="identity")+
      xlab('')+
      ggtitle('Top 20 Sub-Category reviews under Doctors Category')+
      Plot_theme

```

## Next Step

The review content data for each business_id will be analyzed using text mining. A marchine learning method (e.g. SVM, cluster analysis, NNK, etc.) will be employed to build a detection model that would automatically sub-categorize a "Doctors" business.
